Higher Education Synthesized

### 2025-03-15 artifacts/report/msds_data_mining_project_report.tex
adjust conclusion section in report

\section{Conclusion}

    project summary
    key findings
    future work

    This approach is probably over-engineered for the scope of this project, but it lays the groundwork for pulling in much more IPEDS data and running more complicated analyses

    Emphasize where the project currently is at when updating report and slides

    Why synthesize the IPEDS data we use instead of student-level data, the area where we will see a real drop off in coming years? How are aggregates in IPEDS different from student-level data?

    There is not much to conclude at the moment. We will fill this in as we start doing the data work of the project. What I expect to see here is that we trimmed down the IPEDS data, scoped the use case for the project to be a bit narrower, spent more time explore synthetic data and its associated generative modeling, and picked a simpler overall output, likely sticking supervised modeling. For a surprise finish, we may end up deciding to pull some data out and run it through a separate business intelligence tool like Tableau with the idea that end users of the production pipeline may want to leverage tools such as that for their purposes, though this would only come into play if we find that the project turns out to be too quick and simple to get through.

    As for future iterations of the project, I expect that the biggest area of grows would be to incorporate more IPEDS data to answer a broader range of questions. I do not expect radical changes in the pipeline architecture in next iterations of the project, but I do expect additions and expansions in order to accommodate particular questions of interest inherent in the new data.

@@

>>>>

\section{Conclusion}
    To recap, we are using three domains of IPEDS data to test out generative modeling to produce synthetic tabular data, preparing for the reality where we have less and less higher education data in the coming years. We set up a small sample of a full ETL pipeline, send the IPEDS data through, generate synthetic data, and do some exploratory modeling on that data. This approach is probably over-engineered for the scope of this project, but it lays the groundwork for pulling in much more IPEDS data and running more complicated analyses in the future.

    At this point, I am still experimenting with the outputs from the generative modeling. Most of the pipeline and project are complete, and the third iteration of this report and the associated slides will fill in details about modeling on the the synthetic data itself.

    The biggest findings so far have been some of what we expected to see: the IPEDS data contains a lot more complexity than what it looks like at first. We address this by descoping the data that we work with in this pass. In future iterations, we will incorporate much more IPEDS data. That will result in a more robust and complicated warehouse, generative modeling, and exploration of questions with the synthetic data. But, for now, the scope here feels healthy for a first iteration on a project of this sort.

    One area to call out that I do not plan to have an answer for in this project is how we would pivot from IPEDS institutional-level data to student-level data, the latter being the area where are likely to see the biggest drop-off in future years. When doing initial planning for this project, I expected that there would be a couple of obvious areas in which to pivot from IPEDS data to student-level data, but those areas have not revealed themselves. One issue is that student-level data by its nature of being scoped to individual students is not and should not be in large aggregated and publicly available data such as what we find on IPEDS, primarily for privacy concerns. There are many questions that the batch of IPEDS data in this project raise that could lead us to exploring student-level data, but I do not see any reason to go down that path at this point.

    That all said, while generating synthetic IPEDS data is interesting and valuable, it is likely that generating synthetic student-level data may be the real need in the future.

    In terms of the synthetic data itself, while I am still working on that layer of the pipeline and project, so far it appears to work as desired, allowing us to create new data that is similar to the statistical trends we see in the real data. I have pulled in some of my higher education data experience when working with this batch of IPEDS data, but a first next step in a future iteration of the project would be to highlight areas where there are known difficulties and confusion with higher education data and exploring how the generative modeling handles those. And, as said above, the pipeline and approach here could use thorough testing and enhancement related to incorporating a much wider range of IPEDS data, a more complicated warehouse, and improved scope and usage of generative modeling.

@@

    \subsection{Data from Raw Files}
